---
title: "Remote Sensing"
format: html
editor: visual
---

# Detecting Change of Aboveground Biomass using Landsat 9 and PlanetScope data

::: panel-tabset
## Background

Forests play a vital role in regulating the global carbon cycle, yet their aboveground biomass (AGB) remains difficult to estimate over large areas due to limitations of field-based measurements. Remote sensing offers a scalable solution, but the choice of satellite data remains challenging. This study addresses the challenge by evaluating satellite-based approaches to estimate AGB in Kwahu South, Ghana

## Objectives

- Identifying the optimal approach for estimating AGB in Ghana
- Evaluating the accuracy AGB estimation models derived from Landsat 9, PlanetSocpe, and 30-m PlanetScope with GEDI L4A serving as a reference data

## Datasets
- Landsat 9 imagery (date of the images: 12/2022)
- PlanetScope imagery (date of the images: 12/2022)
- GEDI L4A data (date of the data: 09/2022 to 02/2023)
- Digital Elevation Model (DEM)

## Refernce

:::

## Study area

The study area is located in Kwahu South District, Ghana

Total area: 795.76 km2

Main climate: wet semi equatorial 

Forest zone: semi-deciduous 

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Load library
library(leaflet)
library(sf)
library(htmlwidgets)

```


```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide"}
# Load shapefile
kwahu_bound <- st_read("C:/Users/hphung.stu/OneDrive - UBC/Documents/GitHub/hieuphung/files/final_project/AOI_Kwahu_South.shp")

```

