---
title: "Analysis of Suitable Sanctuaries for Humpback Whales"
format: html
editor: visual
---

## Applied Skills

<div style="display: flex; flex-wrap: wrap; gap: 0.5rem; margin-bottom: 1.5rem;">

<span style="background-color: #1f3c73; color: white; padding: 0.5rem 1rem; border-radius: 0.5rem;">ArcGIS Pro</span>

<span style="background-color: #1f3c73; color: white; padding: 0.5rem 1rem; border-radius: 0.5rem;">Suitability Analysis</span>

<span style="background-color: #1f3c73; color: white; padding: 0.5rem 1rem; border-radius: 0.5rem;">Handling Vector Data</span>

</div>

::: panel-tabset
## Background

Hawaii is an archipelago located in the central Pacific Ocean, approximately 3850 km from Californa, USA (World Atlas, n.d). Aside being a wonderful place for visitors, Hawaii archipelago is also home to Humpback Whales during the winter time. The whales swim from Alaska to Hawaii to mate, give birth, and nurture their calves (Hawaii Wildlife Fund, n.d). Recognized as one of the world's most vital humpback whale habitats, the Hawaiian Islands Humpback Whale National Marine Sanctuary was established in 1992 to safeguard humpback whales (Megaptera novaeangliae) and their habitat in Hawai‘i (Division of Aquatic Resources, n.d). However, there have been reports about collisions between whales and human activities and objects such as cruise ships, whale watching boats, and fishing boats. Therefore, government agencies are working to identify the most suitable and expansive areas to incorporate into the sanctuary.

In order to complete the tasks, suitability analysis and overlay analysis were performed. Suitability analysis is a process used to determine the suitability of an area for a particular use or purpose based on given weighted criteria (ESRI a, n.d). Overlay analysis is a common method used in suitability analysis, which combines multiple dataset to create a comprehensive assessment (ArcMap, n,d).

## Objectives

-   The main objectives of this task is to create areas of the 5 most suitable sanctuary for Humpback Whales in Hawaii based on the weighted suitability index.

## Data descriptions

-   The data is obtained from UBC-postgres database. The necessary dataset used in this task are whale species, regions in which the whales appear, human activities that affect the whales.

## References

ArcMap. (n.d). Overlay analysis. https://desktop.arcgis.com/en/arcmap/latest/analyze/commonly-used-tools/overlay-analysis.htm

Division of Aquatic Resources. (n.d). Hawaiian Islands Humpback Whale National Marine Sanctuary (HIHWNMS). https://dlnr.hawaii.gov/dar/marine-managed-areas/hawaiian-islands-humpback-whale-national-marine-sanctuary-hihwnms/

ESRI. (n.d). Perform a suitability analysis. https://pro.arcgis.com/en/pro-app/latest/help/analysis/business-analyst/understanding-suitability-analysis.htm

Hawaii Wildlife Fund. (n.d). Humpback Whale. https://www.wildhawaii.org/wildlife/whales/

World Atlas. (n.d). Maps of Hawaii. https://www.worldatlas.com/maps/united-states/hawaii
:::

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Load library
library(leaflet)
library(sf)
library(htmlwidgets)

```

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide"}
# Load multiple shapefiles 
top_5_sanctuary <- st_read("C:/Users/hphung.stu/OneDrive - UBC/Documents/ArcGIS/Projects/Top_Five_Sanctuary.shp")

sanctuary_buffers <- st_read("C:/Users/hphung.stu/OneDrive - UBC/Documents/ArcGIS/Projects/Sanctuary_Buffer.shp")

most_suitable <- st_read("C:/Users/hphung.stu/OneDrive - UBC/Documents/ArcGIS/Projects/Most_Suitable.shp")

```

```{r, echo = FALSE}
# Create a leaflet map
m <- leaflet() %>% 
  addProviderTiles("Esri.WorldImagery") %>% 
  
  #add sanctuary buffer
  addPolygons(
    data = sanctuary_buffers,
    color = "green",
    weight = 2,
    fillOpacity = 0.5,
    group = "5 Km Buffers") %>% 
  
    #add top_5_sanctuary
  addPolygons(
    data = top_5_sanctuary,
    color = "yellow",
    weight = 2,
    fillOpacity = 0.75,
    group = "Top 5 Sanctuary") %>%
  
  #add most_suitable 
  addPolygons(
    data = most_suitable,
    color = "red",
    weight = 2,
    fillOpacity = 1,
    group = "Most Suitable Area") %>%
  
  #add layer control
  addLayersControl(
    baseGroups = c("Esri World Imagery"),
    overlayGroups = c("Top 5 Sanctuary", "5 Km Buffer", "Most Suitable Area"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  
  #add legend
  addLegend(
    position = "bottomright",
    colors = c("blue", "green", "red"),
    labels = c("Top 5 Sanctuary", "5 Km Buffers", "Most Suitable Area"),
    title = "Legend",
    opacity = 1
  ) %>%
  
  #add scale bar
  addScaleBar(position = c("bottomleft"))

# Save as an interactive HTML file
saveWidget(m, "interactive_map.html", selfcontained = TRUE)
  
```

::: {html}
<iframe src="interactive_map.html" width="100%" height="600px">

</iframe>
:::

The interactive map above shows the top 5 suitable sanctuary for the humpback whales. These areas were identified through a suitability analysis based on three key factors: area size, whale sighting frequency, and vessel traffic. The 5 km buffer zones around each sanctuary highlight potential areas of influence, while the most suitable areas within the sanctuaries are emphasized in red. This visualization helps inform decision-making for marine conservation and habitat protection.


::: {html}
Download Static Map: <a href="files/Lab3_map.pdf" download> <button>Download</button> </a>
:::
