---
title: "Image Classification using R"
format: html
editor: visual
code-fold: true
---

## Applied Skills

::: {style="display: flex; flex-wrap: wrap; gap: 0.5rem; margin-bottom: 1.5rem;"}
[R Coding]{style="background-color: #1f3c73; color: white; padding: 0.5rem 1rem; border-radius: 0.5rem;"}

[QGIS]{style="background-color: #1f3c73; color: white; padding: 0.5rem 1rem; border-radius: 0.5rem;"}

[Supervised Classification]{style="background-color: #1f3c73; color: white; padding: 0.5rem 1rem; border-radius: 0.5rem;"}
:::

::: panel-tabset
## Background

Image classification is a fundamental task in remote sensing used to extract meaningful information from satellite imagery. By assigning each pixel in an image to a specific land cover class (e.g., forest, water, urban, agriculture), classification allows us to analyze spatial patterns, monitor environmental changes over time, and support decision-making in fields like urban planning, agriculture, forestry, and conservation.

## Objectives

-   Classifying the raw Landsat 9 image into different land cover class by using supervised classification method

## Data descriptions

-   Landsat 9 data (including 6 bands)

## References

The instruction for this task was written by the teaching team of MGEM program at The University of British Columbia
:::

The Landsat 9 satellite image shows the Southern Gulf Island, BC. The result of classification can be seen in the swiped map below.

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide"}

library(leaflet)
library(raster)
library(leaflet.extras2)
library(terra)
library(leafem)
library(mapview)

```

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide"}
# Load classified raster (e.g., 2020 classification)
classified <- rast("C:/Users/hphung.stu/Downloads/GEM520/Labs/Lab6/Lab 6 - Classification/Lab 6 - Classification/outputs/classified_map2.tif")
classified_r <- raster(classified)  # convert for leaflet
```

```{r, echo = FALSE}
# Define class colors (hex or names)
landcover_colors <- c(
  "Water" = "violet",
  "Coniferous Forest" = "#d73027",
  "Broadleaf Forest" = "#a50026",
  "Non-Forest Vegetation" = "cyan",
  "Exposed Soil" = "#fdae61",
  "Low Density Developed" = "#91cf60",
  "High Density Developed" = "lightyellow"
)

# Create swipe map with only basemap and classified map
l1 <- leaflet(width = 1200, height = 800) %>%
   setView(lng = -123.1, lat = 48.6, zoom = 11) %>%  # Set center and zoom
  
  # Define left and right panes
  addMapPane("left", zIndex = 2) %>%
  addMapPane("right", zIndex = 1) %>%
  
  # Add basemap to both panes (split for swipe functionality)
  addProviderTiles("Esri.WorldImagery", group = "Basemap-Left", layerId = "baseid_left",
                   options = pathOptions(pane = "left")) %>%
  addProviderTiles("Esri.WorldImagery", group = "Basemap-Right", layerId = "baseid_right",
                   options = pathOptions(pane = "right")) %>%
  
  # Add the classified map to one of the panes (right)
  addRasterImage(classified, group = "Classified Map",
                 maxBytes = 10 * 1024 * 1024,
                 options = pathOptions(pane = "right"), opacity = 1) %>%
  
  # Add custom legend
  addLegend(position = "topright",
            colors = landcover_colors,
            labels = names(landcover_colors),
            title = "Land Cover Classes",
            opacity = 1) %>%
  
  # Swipe control using pane layer IDs
  addSidebyside(
    layerId = "sidecontrols",
    leftId = "baseid_left",
    rightId = "Classified Map",  # show classified over basemap
    options = list(padding = 0)
  ) %>%
  
  addScaleBar(position = "bottomleft")
```

# 

Table 1. Confusion matrix for land cover classification

|   | Broadleaf Forest | Coniferous Forest | Exposed Soil | High Density Developed | Low Density Developed | Non-forest Vegetation | Water |
|---------|---------|---------|---------|---------|---------|---------|---------|
| Broadleaf Forest | 225 | 11 | 0 | 0 | 7 | 0 | 0 |
| Coniferous Forest | 3 | 465 | 0 | 0 | 4 | 0 | 0 |
| Exposed Soil | 0 | 0 | 308 | 2 | 19 | 220 | 0 |
| High Density Developed | 0 | 0 | 28 | 119 | 11 | 7 | 94 |
| Low Density Developed | 4 | 1 | 75 | 40 | 295 | 2 | 1 |
| Non-forest Vegetation | 1 | 0 | 0 | 0 | 9 | 648 | 0 |
| Water | 0 | 0 | 0 | 0 | 0 | 0 | 2821 |

-   Overall accuracy: 0.90

-   Producer's accuracy:

    -   Broadleaf Forest: 0.97

    -   Coniferous Forest: 0.97

    -   Exposed Soil: 0.75

    -   High Density Developed: 0.73

    -   Low Density Developed: 0.86

    -   Non-forest Vegetation: 0.74

    -   Water: 0.97

-   User's accuracy:

    -   Broadleaf Forest: 0.93

    -   Coniferous Forest: 0.99

    -   Exposed Soil: 0.56

    -   High Density Developed: 0.46

    -   Low Density Developed: 0.71

    -   Non-forest Vegetation: 0.98

    -   Water: 1.00
